---
title: "Script3b-H3K27me3_Gene_List"
author: "Abby Deaven"
date: "2025-05-13"
output: html_document
---

```{r setup, include=FALSE}
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")


install.packages("purrr")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("vctrs")
install.packages("ggtext")

library(purrr)
library(tidyverse)
library(vctrs)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library(ComplexHeatmap)
library(InteractiveComplexHeatmap)
library(ggtext)
require("ggrepel")
library(matrixStats)



```

## R Markdown


Reading in important files:
- Matrix of tpm values from RNA-seq samples (Script 2)
- List of H3K27me3 marked genes (Script 3a)
- Metadata

```{r}
all_tpm <- read.table("/MatrixPlotting/Average_Master_TPM.txt", header = T, sep = "\t")

```


```{r Clean up sampleFile and generate final list of H3K27me3-marked genes}
sampleFile <- read.table("./Datasets/RNAseq_Sample_File.txt",skip=0, header=TRUE,stringsAsFactors=FALSE, row.names=2, check.names=FALSE, sep="\t")

K27_K36_ComarkedGenes <- read.table("./Datasets/Final_H3K27me3_H3K36me3_Comarked_NCU.txt", header=FALSE,stringsAsFactors=FALSE, check.names=FALSE, sep="\t")

all_tpm <- read.table("/MatrixPlotting/Average_Master_TPM.txt", header = T, sep = "\t")


## Removing potentially mislabeled samples from sampleFile -- these are mycelial WT samples with very high expression of H3K27me3-marked genes and do not cluster with other samples in correlation plot.
remove_SRRs <- c("SRR13859508", "SRR11266658", "SRR11768109", "SRR5320488")
sampleFile_filtered <- subset(sampleFile, !SRR %in% remove_SRRs)


write.table(sampleFile_filtered, file = "./Datasets/RNAseq_Sample_File.txt", sep = "\t")
###### Filtering list of H3K27me3-marked genes to remove outliers/genes in border regions
#all_tpm <- column_to_rownames(all_tpm, var = "GeneID")

mycelia_SRRs <- subset(sampleFile_filtered, Category == "Controls", select = SRR)
mycelia_tpm <- data.frame(all_tpm[, colnames(all_tpm) %in% mycelia_SRRs$SRR])

#Subset to only include H3K27me3/H3K36me3-marked genes, then calculate median + SD TPM for each gene
mycelia_H3K27me3_tpm <- as.matrix(subset(mycelia_tpm, rownames(mycelia_tpm) %in% K27_K36_ComarkedGenes$V1))
MedianTPM <- data.frame(row.names = rownames(mycelia_H3K27me3_tpm),
                        MedianTPM = rowMedians(mycelia_H3K27me3_tpm))


p <- ggplot(MedianTPM, aes(x = "", y = MedianTPM)) + 
  geom_boxplot(outlier.shape = 2)

## finding statistical outliers
find_outlier <- function(x) {
  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
}


pg <- ggplot_build(p)

MedianTPM_Matrix <- as.matrix(MedianTPM)

MedianTPM_Outliers <- MedianTPM %>% 
  mutate(deepTools_group = ifelse(find_outlier(MedianTPM), "Outlier", "Non_Outlier")   
)

MedianTPM_Outliers <- rownames_to_column(MedianTPM_Outliers, var="GeneID")

p2 <- ggplot(MedianTPM_Outliers, aes(x = outlier, y = MedianTPM)) + 
  geom_boxplot()


### Used deepTools to "generate a heatmap of H3K27me3-marked genes; importing bed file resulting from the first heatmap
Heatmap_Bed <- read.table("./H3K27_H3K36_regions.bed", sep = "\t")
colnames(Heatmap_Bed) = c("chrom", "start", "end", "GeneID", "score", "strand","v7","v8","v9","v10","v11","v12","")


Outlier_bed_heatmap <- left_join(MedianTPM_Outliers, Heatmap_Bed, by = "GeneID")

##removing median tpm column and replacing deepTools-generated clusters
Outlier_Bed_File <- cbind(Outlier_bed_heatmap[,1],Outlier_bed_heatmap[,4:14],Outlier_bed_heatmap[,3])
write.table(Outlier_Bed_File, file = "./H3K27me3_H3K36me3_silent_genes.bed", sep="\t", quote=FALSE)

#Return final set of genes that are not outliers

K27_K36_ComarkedGenes_Silent <- subset(MedianTPM_Outliers, deepTools_group == "Non_Outlier", select = GeneID )
write.table(K27_K36_ComarkedGenes_Silent, file="./Final_Silent_H3K27me3_Genes.txt", sep ="\t")
```

## R Markdown
